---
title: "Intro to Likelihood: Assignment"
output:
  html_document: default
  pdf_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Class_16_Assignment.Rmd") -->

Save this Rmarkdown script as a file named “YourlastName_Assignment16.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_Assignment16.html” and upload the 
resulting file to Canvas to turn it in. 
It is due before our next class.

\

**This homework assignment will test your abilities to:**

1. think critically and creatively about probability distributions

2. extrapolate from in-class exercises to apply likelihood-based inference to arbitrary distributions

3. use your R skillz: function building, data visualization, etc.

\

### Question 1

Choose which probability distribution (from the bank below) would best describe 
the following types of data:

bank: {beta, exponential, gamma, lognormal,
	   normal, uniform, poisson, geometric, 
	   negative binomial, bernoulli, binomial}
	   
1. Lengths of ribbon left over after wrapping presents

- uniform - each length has equal probability of being left over

2. A chess grandmaster's win percentage

- beta distribution - binary outcome, win or lose

3. Number of coding errors (bugs) in student homework assignments

- Poisson distribution - number of times an event occurs within an interval of time

4. Deviation from the mean height in a classroom of toddlers

- normal distribution - height

5. Number of times you have to flip a USB drive over before you successfully put it into your computer

- Geometric distribution - number of trails before first success

6. Number of games before your favorite sports team wins a game

- Geometric distribution - number of trials before first success

7. Human arm length

- Normal distribution - continuous range of human traits

8. Number of seagulls circling a fishing boat

- Poisson distribution - number of times an event occurs within an interval of time


### Question 2

What is the log product rule, and how/why is it used for likelihood-based inference?

- It says that the logarithm of a product can be expressed as the sum of the logarithms of the individual parts.
- It's useful to express products as the logarithm of the product because the probabilities are inherently greater than or equal to one. When you multiple two numbers less than one together, the product will be smaller. This quickly leads to infinitely small numbers that are difficult for the computer to express.


### Question 3

3a) The likelihood of a Poisson-distributed datum ($x$) given 
a value of the rate parameter ($\lambda$) is:
\
$\Large p(x \mid \lambda) = \frac{\lambda^x e^{-\lambda}}{x!}$
\

Write your own function for calculating the log likelihood 
of Poisson-distributed data.  The function should take as its 
arguments the observed data (x) and a value of the rate parameter 
$\lambda$, and it should return the log($p(x \mid \lambda)$).

```{r echo=TRUE}
PoissonDist <- function(x, lambda){
  return(log((lambda^x * exp(-lambda))/factorial(x)))
}
```


3b) Confirm that your function works correctly by calculating the log-likelihood 
of 3 as a draw from a Poisson distribution with a rate parameter of 10, 
and comparing it to the answer you get using the built-in R function `dpois()`.

```{r echo=TRUE}
PoissonDist(3, 10)
log(dpois(3,10))
```


### Question 4

Use maximum likelihood inference to estimate 
the parameter value I used to simulate data from a Poisson distribution, 
which are saved as a .Robj file and linked in the Assignment on Canvas.
You should refer to the lecture .Rmd file for helpful reference 
code that you can cannibalize.

I've included pseudocode below to help you on your journey.
Please carry out each step laid out in the pseudocode.

```{r}
################
#	load in simulated dataset 
#	hint: see the function load()
#		don't forget to specify file paths and use quotes
################

load("/Users/spencerjohnson/Desktop/School25/EEB429/homeworkspace/poisson_data.Robj")

################
# visualize data
#	hint: hist() is always a friend
#	hint: a vertical line on the histogram
#			showing the mean of the sample 
#			is a good sanity check for your inference
################

hist(poisson.data)
abline(v = mean(poisson.data), col = "red", lwd = 2, lty = "dotted")


################
# get MLE:
################

# don't forget that the probability of a shared event 
#	should be calculated as the sum of the logs of the 
#	simple events that make up the shared event
#


# do a grid search over values of lambda
#	extra hint: you can use the histogram to inform the upper limit 
#				of your grid search!

poisson.data
pGuesses <- seq(0,1, by=0.0001) #guesses at the possible values of p
testData <- seq(0,8,by=1)

likelihoods <- log(dbinom(x=2,size=length(poisson.data),prob = pGuesses))
pGuesses[which.max(likelihoods)]

# ML estimate for lambda is:
# 0.002

# plot the likelihood of the data as a function of 
#	each value of lambda you evaluated in your grid search
plot(pGuesses,likelihoods)
abline(v = pGuesses[which.max(likelihoods)], col = "red", lwd = 2, lty = "dotted")
legend("topright",          
       legend="MLE",        
       col="red",           
       lty=2,               
       text.col="black"   
)
# visually indicate which value of lambda is the MLE of the 
#	parameter value used to simulate the data

```