---
title: "Intro to Probability 2: Assignment"
author: "Spencer Johnson"
output:
  html_document: default
  pdf_document: default
language: R
---
<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Class_13_Assignment.Rmd") -->

Save this Rmarkdown script as a file named “YourlastName_Assignment13.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_Assignment13.html” and upload the 
resulting file to Canvas to turn it in. 
It is due before our next class.

\

**This homework assignment will test your abilities to:**

1. think critically and creatively about probability

2. use simulations to build intuition

3. use your R skillz: function building, data visualization, etc.

\

### Question 1

Define the following terms: 

1a) random variable - A variable that is the result of a random event
 
 
1b) probability distribution function - a rule or formula 
that describes the probability of all possible outcomes of a random event
 


### Question 2

<div class="centered">
$\Large {{n}\choose{k}}p^k (1-p)^{n-k}$
</div>

2a) Using your own words, break down the equation above and 

- p - probabilities, n - the number of events in the sequence, k - the number of 'successes'  
- The first part, n choose k calculates the number of possible sequences that satisfy the desired outcome.
describe what each part means (or is doing). 
- p^k calculates the probability of the specific number of successes k in the sequence.
- (1-p)^n-k calculates the probability of **not** successes in the sequence using the opposite of the probability of success (1-p) to the number of **not** successes (n-k).


2b) Using the function of this equation that you coded up in class, calculate the probability of 
seeing 71 successes out of 531 trials, where the probability of success is 0.12345.

```{r, echo = TRUE}
prob <- function(n,k,p){
return(choose(n,k)*p^k*(1-p)^(n-k))
}

prob(531, 71, 0.12345)
```



### Question 3



December through March in the Northern Hemisphere will likely become a high covid prevalence season.
Given the data below (taken from the [Washtenaw Covid Dashboard](https://www.washtenaw.org/3108/Cases)), 
what is the probability that no one in the classroom has an active case of covid?
\

For the purposes of this calculation, assume the following:

 - there are 20 people in our class\
 - each of us can be treated as a random draw from Washtenaw County\
 - there are ~375,000 people in Washtenaw County\
 - over the last week, Washtenaw County has averaged about 29 new cases reported per day\
 - someone infected with covid is infectious for 10 days\
 - only 1 in every 10 covid infections is detected.\

3a) What is the analytical probability that no one in our classroom 
has an active covid infection?
```{r, echo = TRUE}
probCov <- 10 * (29/375000)
probClass <- prob(20, 0, probCov)
probClass
```


\

3b) What is the probability at least one person in the room **_does_** have an active 
covid infection
```{r, echo = TRUE}
probCov <- 10 * (29/375000)
probClass <- prob(20, 0, probCov)
1 - probClass
```

\

3c) What is probability 2 people in the room are infected?
```{r, echo = TRUE}
probCov <- 10 * (29/375000)
probTwoClass <- prob(20, 2, probCov)
probTwoClass
```

\

3d) Write a function for simulating this problem for gatherings of arbitrary size.
The arguments of the function should be the number of people in the room 
and the probability that a randomly selected person in the county has an active covid infection,
and it should return the probability (from simulations), 
that no one in the room has an active covid infection.
Test it with the following class sizes: 10, 41, 100, 500. 

```{r, echo = TRUE}
partyCov <- function(n,probCov = (10 * (29/375000)) ){
  k <- 1
  prob <- choose(n,k)*probCov^k*(1-probCov)^(n-k)
  return(prob)
}

partyCov(10)
partyCov(41)
partyCov(100)
partyCov(500)
```


