---
title: "Class_21_Assignment"
author: "Spencer Johnson"
date: "2025-03-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



################################
################################
# writing an MCMC together
#	by our powers combined
################################
################################

```{r}

# set seed for reproducibility
set.seed(14)

# simulate some data
#	for testing our MCMC
load("/Users/spencerjohnson/Desktop/School25/EEB429/homeworkspace/stamenData.Robj")
head(stamenData)
str(stamenData)

```


- Use a binomial distribution since every group of five is like a sample with a possible 0-5 flowers w stamen
- Creating a prior with a uniform distribution because I'm really not sure what a good prior should be - so I won't use a beta distribution


################
# function library
################


```{r}

# calculate log likelihood of data
#	summed across all observations
calcLik <- function(data, size, prob){
	lik <- sum(dbinom(x=data, size = size, prob = prob, log = TRUE))
	return(lik)	
}

# calculate log prior belief
#	(we don't have strong prior beliefs here)
calcPrior <- function(prob){
	prior <- dunif(x=prob,min=0,max=1,log=TRUE)
	return(prior)	
}

# in log space, this is the sum of the likelihood & prior
calcNumerator <- function(data,size, prob){
	lik <- calcLik(data,size, prob)
	prior <- calcPrior(prob)
	numerator <- lik + prior
	return(numerator)
}

# in log space, this is the difference of the numerator and demoninator
calcRatio <- function(numerator1,numerator2){
	ratio <- numerator1 - numerator2
	return(ratio)
}

# because we're in log space, 
#	 we have to take the exponent of R 
#	(to get out of log-space)
check2accept <- function(R){
	accept <- FALSE
	if(exp(R) > runif(n=1,min=0,max=1)){
		accept <- TRUE
	}
	return(accept)
}
```


################
# Do the MCMC!
################
```{r}

# start with some book-keeping
#	(aka call your mother!)
steps <- 5e3
highprs <- rep(NA,steps)
highpost <- rep(NA,steps)
startProb <- runif(1,0,1)


# initiate the chain for **High Elevation**
pr <- startProb
highprs[1] <- pr
numBayes <- calcNumerator(data = stamenData$highElev, size = 5, prob = pr)
highpost[1] <- numBayes

for(i in 2:steps){
	pr_prime <- pr + rnorm(1,mean=0,sd=0.1)
	if(pr_prime > 0 & pr_prime < 1){
		numBayes_prime <- calcNumerator(data = stamenData$highElev, size = 5, prob = pr_prime)
		R <- calcRatio(numBayes_prime,numBayes)
		if(check2accept(R)){
			pr <- pr_prime
			numBayes <- numBayes_prime
		}
	}
	highprs[i] <- pr
	highpost[i] <- numBayes
}


#######################

lowprs <- rep(NA,steps)
lowpost <- rep(NA,steps)


# initiate the chain for **low Elevation**
pr <- startProb
lowprs[1] <- pr
numBayes <- calcNumerator(data = stamenData$lowElev, size = 5, prob = pr)
lowpost[1] <- numBayes

for(i in 2:steps){
  pr_prime <- pr + rnorm(1,mean=0,sd=0.1)
  if(pr_prime > 0 & pr_prime < 1){
    numBayes_prime <- calcNumerator(data = stamenData$lowElev, size = 5, prob = pr_prime)
    R <- calcRatio(numBayes_prime,numBayes)
    if(check2accept(R)){
      pr <- pr_prime
      numBayes <- numBayes_prime
    }
  }
  lowprs[i] <- pr
  lowpost[i] <- numBayes
}


par(mfrow=c(2,2))
hist(highprs,
     main = "High Elevation", xlab = "Probability")
hist(lowprs,
     main = "Low Elevation", xlab = "Probability")
plot(highprs, highpost, 
     xlab = "Probability", ylab = "Log Posterior Prob.",
     main = "High Elevation")
abline(v = highprs[which.max(highpost)],
       lty = "dotted", col = "red")
plot(lowprs, lowpost,
     xlab = "Probability", ylab = "Log Posterior Prob.",
     main = "Low Elevation")
abline(v = lowprs[which.max(lowpost)],
       lty = "dotted", col = "red")
```


################### Analysis################### 

```{r}
par(mfrow=c(1,2))
plot(highprs, highpost, 
     xlab = "Probability", ylab = "Log Posterior Prob.",
     main = "High Elevation")
abline(v = highprs[which.max(highpost)],
       lty = "dotted", col = "red")
plot(lowprs, lowpost,
     xlab = "Probability", ylab = "Log Posterior Prob.",
     main = "Low Elevation")
abline(v = lowprs[which.max(lowpost)],
       lty = "dotted", col = "red")
```

Based on this set of graphs, the probability of the low elevation and high elevation having stamens if very close. Since we're using the MCMC, I would conclude that there is a slight difference in probability but I would be interested in further analysis to see if the difference is statistically significant.