---
title: "Johnson_Assignment12"
output: html_document
date: "2025-02-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
- probability 
  - A measure of the likelihood that an event will occur
- simple event
  - A single experiment and has a single outcome
- complex event
  - One event made of a series of simple events leading to a final outcome
- shared event
  - The simultaneous occurrence of simple events - the product of the probabilities of it's constitutive simple events 
- conditional probability
  - The probability of one outcome *conditional* on another event
- statistical independence
  - Events occurring independently of other outcomes.


# Question 2 
```{r}
#2a
0.5 * (1/6)

#2b
(4/6) * (1/2)

#2c
(5/6) * 1

#2d
6 * (1/6 * 1/2)

```


# Question 3

- Probability of outcome B is 1 - (p(A)) = 0.8
  - probability of NOT A is 80%


# Question 4
```{r}
Bday <- function(nPeople, replicates = 10000){
pBday <- 0 #sets an internal counter to 0
  for (i in 1:replicates){ # repeats the following sequence for the number of replicates
    
  roomSamp <- sample(1:365, nPeople, replace = TRUE) #Generates a vector of randomly sampled numbers between 1:365, each representing a day of the year
  unique <- duplicated(roomSamp) # Creates a new vector of booleans, if a number had already been listed, it returns true
shared <- sum(unique == FALSE) #Creates an object equal to the sum of false values in the *unique* object vector

if (length(roomSamp) != shared){ # if the there are any true values in the *unique* object vector, the initial vector length and the sum will not be equal. In this case it will add 1 to the counter 
  pBday <- pBday + 1
  }
}
return(pBday/replicates) # Divides the number of times there was a shared birthday by the number of replicates, outputting the percent of time there was a shared value OR the probability that, given the number of people, there will be a shared bday
}

#4b
Bday(23) #Only 23 people reliably give a >50% chance there will be a shared birthday.



#4c
# We can use the same function adjusting the number of days in the intitial sample() function to reflect the number in a martian year - 687

MarsBday <- function(nPeople, replicates = 10000){
pBday <- 0 #sets an internal counter to 0
  for (i in 1:replicates){ # repeats the following sequence for the number of replicates
    
  roomSamp <- sample(1:365, nPeople, replace = TRUE) #Generates a vector of randomly sampled numbers between 1:365, each representing a day of the year
  unique <- duplicated(roomSamp) # Creates a new vector of booleans, if a number had already been listed, it returns true
shared <- sum(unique == FALSE) #Creates an object equal to the sum of false values in the *unique* object vector

if (length(roomSamp) != shared){ # if the there are any true values in the *unique* object vector, the initial vector length and the sum will not be equal. In this case it will add 1 to the counter 
  pBday <- pBday + 1
  }
}
return(pBday/replicates) # Divides the number of times there was a shared birthday by the number of replicates, outputting the percent of time there was a shared value OR the probability that, given the number of people, there will be a shared bday
}
MarsBday(23) # Shockingly, you need the same amount of people for there to be a >50% chance???

```




