---
title: "Johnson_LabAssignment10"
author: "Spencer Johnson"
date: "2025-03-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# A. Developing an intuition for the posterior distribution

- **1.**
  - No matter how many flips are involved or what the true P is equal to, the likelihood and posterior probability are the same.
- **2.**
  - As Nflips increases, the SD of the likelihood distribution increases which is to be expected with a larger sample size.
- **3.**
  - When Nflips = 20, the posterior looks more like the prior. When Nflips = 500, the posterior appears much more similar to the likelihood
- **4.**
  - As you add more data, there is more evidence supporting the likelihood so the prior becomes less significant.
- **5.**
  - When Nflips = 20, the posterior is closer to the true p, as Nflips increases likelihood and the posterior distance from the true p becomes more and more equal, at 500 they're almost the same distance and at 1000 flips they are indistinguishable.
- **6.**
  - When p = 0.1, the likelihood is consistently closer to the true p than the posterior is.
- **7.**
  - When you have a smaller sample size and a "reasonable" p, bayesian inference is good for checking, alongside an MCMC can get you really close to the true P. If you have a larger sample size or "unreasonable" p, using maxmium likekihood inference starts you at a value closer to the actual value of p.
  
# B. Linear models
- **1.**
 - $TimeRunner_{i} \sim N(\mu_{i} = 28.5, \sigma = 3.4)$
```{r}
time <- rnorm(103, mean = 28.5, sd = 3.4)
hist(time)
```
- **2.**
  - $z_{i} = \frac{1}{3.1} x_{i} + 0$
```{r}
pace1 <- time/3.1
plot(pace1~time)
```
- **3.**
  - $y_{i} \sim N(\mu_{i} = f(x_{i}), \sigma = 0.3)$
```{r}
pace2 <- rnorm(103, mean = (1/3.1 * time), sd = 0.3)
plot(pace2~time)
lines(pace1[order(time)] ~ time[order(time)], col = "red", ylim = c(15,40))
```

- **Questions**
  - as the standard deviation increases, the dots get further from the trend line, to be expected, but the slope of the regression also tends to become less as the sd increases. 
  
  - The slope is the per unit (of the independent variable) increase of the response variable. In this case its 1/31 indicating that for every extra minute it takes to finish the race, the pace increases by about 0.323 (1/31) minutes.
  
  - The intercept represents the spot on the y axis where the trend line reaches 0 on the x axis. In this case, at time 0 your pace is also 0 which makes it easy. In other cases if you're sample inherently starts at a non-zero value, you would add that value to the deterministic function.